### 2022-10-18 chapters/conclusion.tex

\paragraph{raising awareness of the energy consumption of the Software}
While this thesis's main focus was to optimize energy consumption using comparative studies, it was easier to say that approach \texttt{X} is greener than approach \texttt{Y}, no matter which metric we were using.
However, it will not be the case for developers when measuring the energy consumption of their programs.
Some tried to give labels, such as (A,B,C\dots), while others translated these raw metrics in an equivalent of fuel consumption (e.g., 1 liter of fuel per 1000 lines of code).
Most of the approaches were to use the carbon emission, such as~\cite{patterson2021carbon} or ecograder~\link{https://ecograder.com/} as a pivot metrics to quantify the environmental cost of osftware.
We intend to explore these approaches and study how they can be used to increase awareness of the energy consumption of their code.

\subsection*{Representativeness}:
As reported in the state of the art, a successful benchmark meets three criteria: reproducibility, accuracy, and representativeness.
This Ph.D. thesis mostly discussed two of these criteria: reproducibility, and accuracy.
However, we did not discuss the representativeness of our benchmarks.
We relied on state-of-the-art benchmarks to represent real-world applications.
Nevertheless, the gap between these benchmarks and the industry is getting bigger and bigger, due to the extreme pace of software development.
In the future, we would like to tackle the issue of representativeness and discuss how we can improve our protocol to fit real-world applications.
After all, what is the purpose of doing optimizations if they cannot be applied to real-world applications?

>>>>

\paragraph{Raising awareness about the energy consumption of software}
As the core focus of this PhD thesis was to optimize energy consumption using comparative studies, it was straightforward to claim that approach \texttt{X} is greener than \texttt{Y}, no matter the metrics we were using.
However, this cannot be the case for developers when measuring the energy consumption of their programs.
Some tried to give labels, such as (A,B,C\dots), while others translated these raw metrics in an equivalent of fuel consumption (e.g., 1 liter of fuel per $1,000$ lines of code).
Most of the approaches were to use the carbon emission, such as~\cite{patterson2021carbon} or ecograder,\link{https://ecograder.com/} as a pivot metrics to quantify the environmental cost of software.
We intend to explore these approaches and study how they can be used to increase awareness of the energy consumption of their code.

\subsection*{Representativeness}
As reported in the state of the art, a successful benchmark meets three criteria: reproducibility, accuracy, and representativeness.
This Ph.D. thesis mostly discussed two of these criteria: reproducibility, and accuracy.
However, we did not discuss the representativeness of our benchmarks.
We relied on state-of-the-art benchmarks to represent real-world applications.
Nevertheless, the gap between these benchmarks and the industry is getting bigger and bigger, due to the extreme pace of software development.
In the future, we would like to tackle the issue of representativeness and discuss how we can improve our protocol to fit real-world applications.
After all, what is the purpose of doing optimizations if they cannot be applied to real-world applications?

### 2022-10-18 chapters/optimization/python-dir/python.tex

Figure~\ref{fig:python_multiprocessing} shows the results of the experiment regarding four metrics: power, execution time, DRAM energy, and CPU energy.
As one can notice, CPU energy is ten folds higher than the DRAM one.
Therefore, we will focus on the CPU for this study.

@@

\paragraph*{Synthesis}
This study demonstrates that the optimal way to reduce the energy consumption of Python applications is to adopt the \textsc{Multiprocessing} strategy, which is the most energy-efficient one.
However, this strategy should be used carefully as it can lead to increased execution times.
This study demonstrates that the optimal number of processes is equal to the number of physical cores, which is four in our case.
When using the \textsf{Multithreading} strategy, it was also found that sometimes the performance of the application has to be given up to save energy.

@@

In the second section, we presented a non-intrusive technique to optimize the energy consumption of Python-based apps without making substantial changes to the code.
This technique consists in guiding the choice of the Python runtime implementation.
We started by categorizing and filtering these implementations into three significant classes (compiler, interpreter, and extra libraries).
Then, we conducted a series of experiments to compare the energy consumption of these alternatives.
Our findings indicate the lack of a general solution and the importance of tuning the Python runtime depending on applications.
% In some circumstances, some of these answers may be the best, while others may be the worst.
We found that most interpreters had a similar or worse energy consumption than the official implementation of CPython.
The reason behind such behavior was that some implementations focused on specific case studies, such as machine learning or security.
In contrast, others focused on the compatibility of the Python code with their platforms, like Jython and IronPython.
Finally, regardless of implementation, we demonstrated that using JIT is the most efficient technique to boost the energy efficiency of Python-based programs.

>>>>

Figure~\ref{fig:python_multiprocessing} shows the results of the experiment regarding four metrics: power, execution time, DRAM energy, and CPU energy.
As one can notice, CPU energy is ten folds higher than DRAM one.
Therefore, we will focus on the CPU for this study.

@@

\paragraph*{Synthesis}
This study demonstrates that the optimal way to reduce the energy consumption of Python applications is to adopt the \textsc{Multiprocessing} strategy, which is the most energy-efficient one.
However, this strategy should be used carefully as it can lead to increase power consumption.
This study demonstrates that the optimal number of processes is equal to the number of physical cores, which is four in our case.
When using the \textsf{Multithreading} strategy, it was also found that sometimes the performance of the application has to be given up to save energy.

@@

In the second section, we presented a non-intrusive technique to optimize the energy consumption of Python-based apps without making substantial changes to the code.
This technique consists in guiding the choice of the Python runtime implementation.
We started by categorizing and filtering these implementations into three significant classes (compiler, interpreter, and extra libraries).
Then, we conducted a series of experiments to compare the energy consumption of these alternatives.
Our findings indicate the lack of a general solution and the importance of tuning the Python runtime depending on the application.
% In some circumstances, some of these answers may be the best, while others may be the worst.
We found that most interpreters had a similar or worse energy consumption than the official implementation of CPython.
The reason behind such behavior was that some implementations focused on specific case studies, such as machine learning or security.
In contrast, others focused on the compatibility of the Python code with their platforms, like Jython and IronPython.
Finally, regardless of implementation, we demonstrated that using JIT is the most efficient technique to boost the energy efficiency of Python-based programs.
